---
title: "Assignment_2"
author: "Harsha Singh"
date: "27/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r cars}
summary(cars)
```

Assignment - 2
1.Result of final world cup of football is France- 4 and Croatia – 2. Please use the control structure to print results as Team ---- Wins. ----- should be country name.
```{r}
Result<-c(4,2)
names(Result)<-c("France", "Croatia")
Result
if (Result["France"] > Result["Croatia"]){
  print("Team France Wins")
} else if(Result[France]< Result["Croatia"]){
  print("Team Croatia wins")
}
```
```{r}
France <- 4
Croatia <- 2

if (France < Croatia){
  print("Team Croatia Wins")
}  else {
    print("Team France Wins")
}
getwd()
```


mtcars dataset has several factor variables. However, R is reading them numeric. Please convert them into factor using a for loop. Please use column 8 to 11 for loop.
```{r}
str(mtcars)
for (row in (8:11)){
  if(class(mtcars[[row]])!="row"){
  mtcars[[row]]<- as.factor(mtcars[[row]])
  }
}
```


```{r}
x=0
for(i in 1:dim(mtcars)[2])
{
 if(class(mtcars[,i])=="numeric") {x=x+1}
}
x
str(mtcars)

```


```{r}
for (i in 8:11){
  mtcars[, i] <- as.factor(mtcars[, i])
}
#check if converted
```


```{r}
x=0
for(i in 1:dim(mtcars)[2])
{
 if(class(mtcars[,i])=="factor") {x=x+1}
}
x
factor_data <-factor(mtcars)
print(factor_data)
data(mtcars)

is.factor(mtcars$gear)
```


```{r}
str(mtcars)
```

Here is the Dataset 2.1 for the following question. This dataset is a subset of a real dataset.
Let's read the data-set
```{r}
install.packages(readxl)
install.packages('DT')
####The main function in this package is datatable(). It creates an HTML widget to display R data objects with DataTables.


library(DT)
library(readxl)
vehicles <- read_excel("Dataset - 2.1.xlsx")
datatable(vehicles)
```



Write a function to get percentage of NAs in each column.
```{r}
count<-apply(vehicles, 2, function(x) sum(is.na(x)))
dims<-dim(vehicles)
dims
dims[1]
num<-dims[1]*dims[2]
NApercentage<-(count/num) * 100
print(NApercentage)
```

Write a function to get percentage of NAs in each row.
```{r}
count<-apply(vehicles, 1, function(x) sum(is.na(x)))
dims<-dim(vehicles)
dims
dims[1]
num<-dims[1]*dims[2]
NApercentage<-(count/num) * 100
print(NApercentage)
```


```{r}
colSums(is.na(vehicles))
```


```{r}
# writing function to create percentage
# First step is to write a pseudo code to calculate percentage
# (sum(is.na(x))/length(x))*100
pMiss <- function(x){
  (sum(is.na(x))/length(x))*100
}

missing_perc_col <- apply(vehicles, 2, pMiss)
head(missing_perc_col)
missing_perc_col
```


```{r}
pMiss <- function(x){
  (sum(is.na(x))/length(x))*100
}

missing_perc_row <- apply(vehicles, 1, pMiss)
head(missing_perc_row)
```

Write a function to get summary of numeric columns (use summary function) such as THC, CO, CO2 etc. in data set. In same function try to generate boxplot using base R.
```{r}
find_summary<-function(column){
  print(summary(column))
  boxplot(column)
}

for(col in 1:ncol(vehicles)){
  if(class(vehicles[[col]])=="numeric"){
    find_summary(vehicles[[col]])
  }
}
```

Write a function to create histograms of numeric columns such as THC, CO, CO2 etc. in data set. Use ggpot2 to generate figure.
```{r}
install.packages("ggplot2")
library(ggplot2)

generator<- function(column){
  ggplot(data= vehicles, aes(column),bins = 10) + 
    geom_histogram()
}
generator(vehicles[[1]])
for(col_num in 1:ncol(vehicles)){
  if(class(vehicles[[col_num]])=="numeric"){
    print(generator(vehicles[[col_num]]))
  }
}
```


```{r}
data_dis <- function(x, y){
  ggplot(x) +
    aes(y) +
    geom_histogram() 
}

data_dis(vehicles, vehicles$CO2)
```

Here is the Dataset 2.2 for the following question. This dataset is a subset of a real dataset.
```{r}
library(readr)
FD_data = read_csv("Data set - 2.2.csv")
str(FD_data)
```

Data set contains date columns. All of these date columns are untidy. Please create a better formatted dataset. Date should be dd/mm/yyyy in final format. Use column 2, 3 and 5 only
```{r}
##Observe 2,3,5 are character columns

library(lubridate)
FD_data$"First FD Date"<-mdy(FD_data$"First FD Date")
format(FD_data$"First FD Date","%d/%m/%Y")

FD_data$"Last FD Date"<-mdy(FD_data$"Last FD Date")
format(FD_data$"Last FD Date","%d/%m/%Y")

FD_data$"FD termination date"<-mdy(FD_data$"FD termination date")
format(FD_data$"FD termination date","%d/%m/%Y")

```

Date of birth column contains month in string format. Please create a tidy data column with months in numeric format. Now your data should be similar to previous question.
```{r}
FD_data$"Date of Birth"<-dmy(FD_data$"Date of Birth")
format(FD_data$"Date of Birth","%/%m/%Y")
date_of_birth = dmy(FD_data$`Date of Birth`, locale = 'english')
date_of_birth<-as.Date(FD_data$"Date of Birth", format="%d/%m/%Y")

date_of_birth
date_of_birth <- as.Date(date_of_birth, format= '%d/%m/%y')
date_of_birth = as.Date(format(date_of_birth, '19%y%m%d'), '%Y%m%d')

date_of_birth
FD_data$`Date of Birth` <- date(as.numeric(FD_data$`Date of Birth`, "%d%m%y"))
FD_data$`Date of Birth`
```

Convert all dates into Date format; they are currently in character variable format.
Create a new column age based on date of birth column and First FD column. Date format is necessary to do basic arithmetic.
```{r}
myFun <- function(x, dummyDay = "01", dummyYear = "2013"){
  require(lubridate)
  x <- ifelse(substr(x, 1, 3) %in% month.abb,
              paste(match(substr(x, 1, 3), month.abb),
                    dummyDay,
                    dummyYear, sep = "/"), x)
  #return date
  mdy(x)
}

res <- data.frame(lapply(FD_data$`Date of Birth`, myFun))
```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
